---
title: "Data Science and Business Analytics Internship - The Sparks Foundation"
author: "Shivam Agrawal"
date: "06/10/2021"
output: 
  html_document:
    theme: yeti
    highlight: zenburn
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 999)
```

<br>

# #GRIPOCT2021

<br>

# Introduction

<font size = "4"> Hi my name is Shivam Agrawal. This is my second task. I am using **R Programming** and **RStudio** for this task.</font>

<br>

# TASK-3: Exploratory Data Analysis - Retail

<font size = "4"> **Problem Statement: ** As a business manager, try to find out the weak areas where you can work to make more profit. <br>
**Dataset link: ** https://bit.ly/3i4rbWl <br>

<br>

## Importing libraries

```{r message=FALSE, warning=F}
library(dplyr) #efficiently manipulating datasets in R
library(ggplot2) #For visualization
library(plotly) #For interactive visualization
library(DT) # 
library(RColorBrewer)
```

<br>

## Reading data

```{r message=FALSE, warning=F}
supermarket_data <- read.csv(file = "SampleSuperstore.csv", header = T, stringsAsFactors = F)
datatable(head(supermarket_data, 10), options = list(
                          scrollX = TRUE
                      ))
```

<br>

## Data Analysis

### Structure of data

```{r message=FALSE, warning=F}
glimpse(supermarket_data)
```

### Missing Data?

```{r message=FALSE, warning=F}
supermarket_data %>%
  select(everything()) %>%
  summarise_all(funs(sum(is.na(.)))) %>%
  t()
```

### Dropping column "Postal.Code"

```{r message=FALSE, warning=F}
supermarket_data <- supermarket_data %>%
  select(-c(Postal.Code))
```

### Let's check summary statistics

```{r message=FALSE, warning=F}
supermarket_data %>%
  select_if(is.numeric) %>%
  summary()
```

## Visualization

### Which Shipping Mode mostly preferred?

```{r message=FALSE, warning=F}
df <- supermarket_data %>%
  group_by(Ship.Mode) %>%
  summarise(count = n())  
df$Ship.Mode <- factor(df$Ship.Mode, levels = unique(df$Ship.Mode)[order(df$count, decreasing = TRUE)])

df %>%
  plot_ly(x = ~Ship.Mode,
          y = ~count,
          type = "bar",
          color = ~Ship.Mode,
          hovertemplate = paste("<b>%{x}: %{y}</b><br><br><extra></extra>")) %>%
  layout(title = "Shipping Mode VS Count",
         xaxis = list(title = "Shipping Mode"),
         yaxis = list(title= "Count"))
```

**Observations: ** Standard Class is the most preferred shipping mode.

Does **Standard Class** has higher sales and profits as well ?

```{r message=FALSE, warning=F}
datatable(supermarket_data %>%
  select(Ship.Mode, Sales, Profit) %>%
  group_by(Ship.Mode) %>%
  summarise(Sales = sum(Sales), Profit = sum(Profit)) %>%
  arrange(desc(Sales), desc(Profit)))
```

<br>

### Segment 

```{r message=FALSE, warning=F}
df <- supermarket_data %>%
  group_by(Segment) %>%
  summarise(count = n())  
df$Segment <- factor(df$Segment, levels = unique(df$Segment)[order(df$count, decreasing = TRUE)])

df %>%
  plot_ly(x = ~Segment,
          y = ~count,
          type = "bar",
          color = ~Segment,
          hovertemplate = paste("<b>%{x}: %{y}</b><br><br><extra></extra>")) %>%
  layout(xaxis = list(title = "Segment"),
         yaxis = list(title= "Count"),
         title = "Segments VS Count")
```

**Observations: ** Consumer Segment is overpowering.

Does **Consumer Segment** has higher sales and profits as well ?

```{r message=FALSE, warning=F}
datatable(supermarket_data %>%
  select(Segment, Sales, Profit) %>%
  group_by(Segment) %>%
  summarise(Sales = sum(Sales), Profit = sum(Profit)) %>%
  arrange(desc(Sales), desc(Profit)))
```

<br>

### Categories

```{r message=FALSE, warning=F}
df <- supermarket_data %>%
  group_by(Category) %>%
  summarise(count = n())  
df$Category <- factor(df$Category, levels = unique(df$Category)[order(df$count, decreasing = TRUE)])

df %>%
  plot_ly(x = ~Category,
          y = ~count,
          type = "bar",
          color = ~Category,
          hovertemplate = paste("<b>%{x}: %{y}</b><br><br><extra></extra>")) %>%
  layout(xaxis = list(title = "Category"),
         yaxis = list(title= "Count"),
         title = "Categories VS Count")
```

**Observations: ** Office Supplies are more preferred then other categories.

Does **Office Supplies** has higher sales and profits as well ?

```{r message=FALSE, warning=F}
datatable(supermarket_data %>%
  select(Category, Sales, Profit) %>%
  group_by(Category) %>%
  summarise(Sales = sum(Sales), Profit = sum(Profit)) %>%
  arrange(desc(Sales)))
```

<br>

### Sub-Categories

```{r message=FALSE, warning=F}
df <- supermarket_data %>%
  group_by(Sub.Category) %>%
  summarise(count = n())  
df$Sub.Category <- factor(df$Sub.Category, 
                          levels = unique(df$Sub.Category)[order(df$count, decreasing = TRUE)])

df %>%
  plot_ly(x = ~Sub.Category,
          y = ~count,
          type = "bar",
          color = ~Sub.Category,
          colors = c("darkgreen", "blue", "red"),
          hovertemplate = paste("<b>%{x}: %{y}</b><br><br><extra></extra>")) %>%
  layout(xaxis = list(title = "Sub Categories"),
         yaxis = list(title= "Count"),
         title = "Sub Categories VS Count")
```

**Observations: ** Binders are the most popular sub category.

Does **Binders** has higher sales and profits as well ?

```{r message=FALSE, warning=F}
datatable(supermarket_data %>%
  select(Sub.Category, Sales, Profit) %>%
  group_by(Sub.Category) %>%
  summarise(Sales = sum(Sales), Profit = sum(Profit)) %>%
  arrange(desc(Sales)))
```

<br>

### Top States

```{r message=FALSE, warning=FALSE}
df <- supermarket_data %>%
  group_by(State) %>%
  summarise(count = n()) %>%
  top_n(10) 
df$State <- factor(df$State, 
                          levels = unique(df$State)[order(df$count, decreasing = TRUE)])


df %>%
  plot_ly(x = ~State,
          y = ~count,
          type = "bar",
          color = ~State,
          colors = c("darkgreen", "blue", "red"),
          hovertemplate = paste("<b>%{x}: %{y}</b><br><br><extra></extra>")) %>%
  layout(title = "Top 10 States with maximum Retail",
         xaxis = list(title = "State"),
         yaxis = list(title= "Count"))
```

**Observations: ** California is the state which has most retail 

<br>

### Top Cties

```{r message=FALSE, warning=FALSE}
df <- supermarket_data %>%
  group_by(City) %>%
  summarise(count = n()) %>%
  top_n(10) 
df$City <- factor(df$City, 
                          levels = unique(df$City)[order(df$count, decreasing = TRUE)])


df %>%
  plot_ly(x = ~City,
          y = ~count,
          type = "bar",
          color = ~City,
          colors = c("darkgreen", "blue", "red"),
          hovertemplate = paste("<b>%{x}: %{y}</b><br><br><extra></extra>")) %>%
  layout(title = "Top 10 Cities with maximum Retail",
         xaxis = list(title = "City"),
         yaxis = list(title= "Count"))
```

**Observations: ** New York City is the city which has most retails

<br>

### Region wise Retail

```{r message=FALSE, warning=F}
df <- supermarket_data %>%
  group_by(Region) %>%
  summarise(count = n())

df %>%
  plot_ly(labels = ~Region,
                 values = ~count,
                 marker = list(colors = ~Region)) %>%
         add_pie(hole = 0.2) %>%
         layout(title = "Region wise Retai",
                xaxis = list(zeroline = F,
                             showline = F,
                             showticklabels = F,
                             showgrid = F),
                yaxis = list(zeroline = F,
                             showline = F,
                             showticklabels=F,
                             showgrid=F))
```

**Observations: ** From above plots its clear that West Region has most Retail then other Regions

<br>

### Discount Vs Profit for all Region

```{r message=FALSE, warning=F}
df <- supermarket_data %>%
  select(Profit, Discount, Region) %>%
  group_by(Discount, Region) %>%
  summarise(profit = mean(Profit), .groups = 'drop')

df %>%
 plot_ly(x=~Discount) %>%
         add_markers(y = ~profit,
                     text = ~paste("Profit: ", profit),
                     showlegend = F) %>%
         add_lines(y = ~fitted(loess(profit ~ Discount)),
                   name = ~Region,
                   color = ~Region,
                   showlegend = T,
                   line = list(width=5))  %>% 
  layout(title = "Discount VS Profit",
         xaxis = list(title = "Discount",
                      
                      zeroline = FALSE),
         yaxis = list(title = "Profit",
                     
                      zeroline = FALSE))
```

**Observations: ** From above plot we can see that for higher discounts profit is less in every Regions.

<br>

### Region wise Sales/Profit

```{r message=FALSE, warning=F}
df <- supermarket_data %>%
  select(Region, Profit, Sales) %>%
  group_by(Region) %>%
  summarise(sales = sum(Sales), profit = sum(Profit), count = n())

df$Region <- factor(df$Region, 
                          levels = unique(df$Region)[order(df$count, decreasing = TRUE)])

p1 <- df %>% 
  plot_ly(x = ~Region, y = ~sales, name = "Sales", type = "bar")

p1 <- p1 %>% add_trace(y = ~profit, name = "Profit") %>%
  layout(title = "Region wise Sales/Profit",
         yaxis = list(title = "Sales/Profit"),
         xaxis = list(title = "Region"),
         barmode = "group")
p1
```

**Observations: **

<br>

### State by Sales/Profit

```{r message=FALSE, warning=F}
df <- supermarket_data %>%
  select(State, Profit, Sales) %>%
  group_by(State) %>%
  summarise(sales = sum(Sales), profit = sum(Profit), count = n()) %>%
  top_n(10)

df$State <- factor(df$State, 
                          levels = unique(df$State)[order(df$sales, decreasing = TRUE)])

p1 <- df %>% 
  plot_ly(x = ~State, y = ~sales, name = "Sales", type = "bar")

p1 <- p1 %>% add_trace(y = ~profit, name = "Profit") %>%
  layout(barmode = "group",
         title = "State wise Sales/Profit",
         yaxis = list(title = "Sales/Profit"),
         xaxis = list(title = "State"))
p1

```

**Observations: ** California and New York are the top two states in US for getting highest sales and profit

<br>

### City by Sales/Profit

```{r message=FALSE, warning=F}
df <- supermarket_data %>%
  select(City, Profit, Sales) %>%
  group_by(City) %>%
  summarise(sales = sum(Sales), profit = sum(Profit), count = n()) %>%
  top_n(10)

df$City <- factor(df$City, 
                          levels = unique(df$City)[order(df$sales, decreasing = TRUE)])

p1 <- df %>% 
  plot_ly(x = ~City, y = ~sales, name = "Sales", type = "bar")

p1 <- p1 %>% add_trace(y = ~profit, name = "Profit") %>%
  layout(barmode = "group",
         title = "City wise Sales/Profit",
         yaxis = list(title = "Sales/Profit"),
         xaxis = list(title = "City"))
p1
```

**Observations: ** NYC and Los Angeles are the top two Cities in US for getting highest sales and profit

<br>

### **Conclusion**

From above we can conclude that: <br>

1. **Standard Class** is the most preferred shipping mode and it has also contributed **highest** sales and profit. <br>
2. **Consumer** is the most preferred segment and it has also contributed **highest** sales and profit. <br>
3. **Office Supplies** are the most demanded category but it has contributed **lowest** sales and profit. <br>
4. **Binders** are the most demanded sub category but it has not contributed much. Instead **Phones** has contributed **highest** sales and profit. <br>
5. **California and New York** are the top most States in US in terms of retail and also contributed **highest** sales and profit. <br>
6. **New York City and Los Angeles** are the top most cities in US in terms of retail and also contributed **highest** sales and profit. <br>
7. **West** Region contributed **maximum** retails and also **highest** contributer in terms of sales and profit. <br><br>

Also:<br>

1. We have to work on **discounts** to increase the profit especially in **Central** region.
2. To increase sales in **South** region we have to introduce new ways of sales and marketing.
3. Many states like **Texas**, **Florida** etc. has negative profits. We have to investigate further why these state have negative profit.
4. We have to work on **Corporate** segment as it has least profit then other two segments.
5. **Office Supplies** has the highest retails but it has least sales and higher profit. We have to check the cause and need to work on it to improve the sales.
6. To get more sales and profit in **NYC and Los Angeles** we can introduce different offers.

<br>

### End of Task-3

<br>