---
title: "Data Science and Business Analytics Internship - The Sparks Foundation"
author: "Shivam Agrawal"
date: "03/10/2021"
output: 
  html_document:
    theme: yeti
    highlight: zenburn
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 999)
```

<br>

# #GRIPOCT2021

<br>

# Introduction

<font size = "4"> Hi my name is Shivam Agrawal. This is my first task. I am using **R Programming** and **RStudio** for this task.</font>

<br>

# TASK-1: Prediction using Supervised ML

<font size = "4"> **Problem Statement: ** Predict the percentage of an student based on the number of study hours. <br>
**Model used: ** Simple Linear Regression <br>
**Dataset link: ** http://bit.ly/w-data <br>

<br>

## Importing libraries

```{r message=FALSE}
library(dplyr) #efficiently manipulating datasets in R
library(ggplot2) #For visualization
library(gensvm) #For splitting the data
library(Metrics) #For evaluating the model
```

<br>

## Reading data

```{r}
url <- "http://bit.ly/w-data"
student_data <- read.csv(file = url)
print("Data imported successfully!!")
head(student_data, 10)
```

<br>

```{r}
#Data information
glimpse(student_data)
```

<font size = "3"> Our data contains **25 rows and 2 columns** respectively. </font>

<br>

```{r}
#Descriptive Statistics
summary(student_data)
```

<br>

## Missing values?

```{r warning=FALSE}
student_data %>%
  select(everything()) %>%
  summarise_all(funs(sum(is.na(.)))) %>%
  t()
```

<font size = "3">**No** missing values in our data. </font>

<br>

## Visualization

```{r}
#Our custom theme
mytheme <- theme(plot.background = element_rect(fill="#F5FFFA", color = "darkblue"),
             plot.title = element_text(size=20, hjust = 0.5),
             axis.title.x = element_text(size=18, color = "black"),
              axis.title.y = element_text(size=18, color = "black"),
             axis.text.x = element_text(size=15),
             axis.text.y = element_text(size=15),
            legend.text = element_text(size=15),
            legend.title = element_text(size=15))
```

<font size = "3"> Let's plot the data to see if there is any relationship between the data or not </font>

```{r}
#Scatter Plot
ggplot(data = student_data,
       aes(x = Hours,
           y = Scores)) +
  geom_point(color = "blue", size = 2) + 
  scale_x_continuous(breaks = seq(1,10,1)) +
  scale_y_continuous(breaks = seq(10, 100, 10)) +
  xlab("Hours Studied") +
  ylab("Percentage Score") +
  ggtitle("Hours vs Scores") +
  mytheme
```

<font size = "3"> We can see from the above plot that, there is a **positive linear relationship** between Hours Studied and Percentage Score. </font>


<font size = "3"> Let's see the distribution plot </font>

```{r}
#Hours
ggplot(data = student_data,
       aes(x = Hours)) +
  geom_histogram(aes(y = ..density..),
                 bins = 6,
                 colour = "black",
                 fill="light blue") +
  geom_density(alpha = 0.2, 
               fill = "#FF6666") +
  scale_x_continuous(breaks = seq(1,10,1)) +
  xlab("Hours Studied") +
  ylab("Density") +
  ggtitle("Distribution Plot for Hours Studied") +
  mytheme
```

```{r}
#Scores
ggplot(data = student_data,
       aes(x = Scores)) +
  geom_histogram(aes(y = ..density..),
                 bins = 6,
                 colour = "black",
                 fill="light blue") +
  geom_density(alpha = 0.2, 
               fill = "#FF6666") +
  scale_x_continuous(breaks = seq(10,100,10)) +
  xlab("Percentage Score") +
  ylab("Density") +
  ggtitle("Distribution Plot for Percentage Score") +
  mytheme
```

<br>

## Preparing the Data

<font size = "3"> Now we are splitting data into training and testing sets

```{r}
split_data <- gensvm.train.test.split(x = student_data, train.size = 0.8, shuffle = T, random.state = 0)
```

```{r}
train_set <- split_data$x.train
test_set <- split_data$x.test
```

```{r}
#View train and test sets
train_set
test_set
```

<br>

## Training the Algorithm

```{r}
#Linear Regression Model
model <- lm(Scores ~ Hours, data = train_set)
```

<font size = "3"> Let's see the summary of our linear regression model </font>

```{r}
summary(model)
```

<br>

## Plotting the Regression line

```{r}
ggplot(data = student_data,
       aes(x = Hours,
           y = Scores)) +
  geom_point(color = "dark green", size = 2) + 
  geom_smooth(method = "lm", fill = NA) +
  scale_x_continuous(breaks = seq(1,10,1)) +
  scale_y_continuous(breaks = seq(10, 100, 10)) +
  xlab("Hours Studied") +
  ylab("Percentage Score") +
  ggtitle("Hours vs Scores") +
  mytheme
```

<br>

## Making Predictions

<font size = "3"> Let's make some predictions on test set </font>

```{r}
test_hours <- data.frame(Hours = test_set$Hours)
predicted_scores <- predict(model, test_hours)
```

```{r}
#Compare the actual and predicted scores
df <- data.frame(Hours = test_hours, Actual_Scores = test_set$Scores, Predicted_Scores = predicted_scores)
df
```

<br>

## What will be the predicted score if a student studies for 9.25 hrs/day?

```{r}
hours <- 9.25
new_data <- data.frame(Hours = hours)
stud_pred <- predict(model, new_data)

print(paste0("If a student studies for ", hours, " hrs/day", " then he will score (predicted): ",
             stud_pred))
```

<br>

## Evaluating the Model

<font size = "3"> Let's evaluate the performance of the model. This step is particularly important to compare how well different algorithms perform on a particular dataset. We will use **Mean Absolute Error, Mean Squared Error and Root Mean Squared Error**. </font>

```{r}
print(paste0("Mean Absolute Error: ", mae(test_set$Scores, predicted_scores)))
print(paste0("Mean Squared Error: ", mse(test_set$Scores, predicted_scores)))
print(paste0("Root Mean Squared Error: ", rmse(test_set$Scores, predicted_scores)))
```

<br>

## Accuracy of the Model

```{r}
summary_model <- summary(model)
summary_model$r.squared
```

<br>

## End of Task-1

<br>